An exposition on the propriety of restricted Boltzmann machines

### 2016-09-03 writing/draft.Rmd
update beginning of exporations section with Vardeman's comments

We explore the relationship between magnitudes of $\boldsymbol \theta$ and the three measures of model impropriety, near-degeneracy, instability, and uninterpretability, for varying sized RBM models through simulated data.

## Tiny example
To illustrate the ideas of graph near-degeneracy, instability, and uninterpretability in a RBM, we will consider a the smallest possible toy example that consists of one visible node $v_1$ and one hidden node $h_1$, both binary. A schematic of this model can be found in figure \ref{fig:toymodel}.

\begin{figure}[ht]
  \centering
  \resizebox{1cm}{!}{\input{tikz/toymodel.tikz}}
  \caption{A small example restricted Boltzmann machine (RBM), which consists of two nodes, one hidden and one visible.}
  \label{fig:toymodel}
\end{figure}

@@

Additionally, we can compute the mean parametrization of the model parameters as 
\begin{align*}
\mu(\boldsymbol \theta) &= \text{E}_{\boldsymbol \theta}\left[ t(X) \right] \\
&= \sum\limits_{x \in \mathcal{X}} \left\{ t(x) \frac{\exp\left( \theta_{11} h_1 v_1 + \theta_{h1} h_1 + \theta_{v1} v_1\right)}{\sum\limits_{\boldsymbol x \in \mathcal{X}}\exp\left( \theta_{11} h_1 v_1 + \theta_{h1} h_1 + \theta_{v1} v_1\right)} \right\} \\
&= \sum\limits_{x \in \mathcal{X}} \left\{ t(x) \frac{\exp\left( \theta_{11} h_1 v_1 + \theta_{h1} h_1 + \theta_{v1} v_1\right)}{\gamma(\boldsymbol \theta)} \right\} \\
&= \left[
\begin{matrix}
\frac{\exp\left(\theta_{v_1}\right) + \exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\frac{\exp\left(\theta_{h_1}\right) + \exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\frac{\exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\end{matrix} \right]
\end{align*}
These three functions can be visualized over multiple values of $\boldsymbol \theta$, as seen in figure \ref{fig:toymodel}. By examining these functions, we can see that as values of $\boldsymbol \theta$ grow larger in magnitude, at least one mean-parameterization functions is approaching the values 0 or 1, indicating a potential to be near to the boundary of the convex hull of $\{t_k(\boldsymbol x): \boldsymbol x \in \mathcal{X}, k = 1, \dots, K\}$. Thus, for a very small example we can see the relationship between values of $\boldsymbol \theta$ and model degeneracy.

```{r near_dependency, fig.cap="\\label{fig:toymodel}The three mean-parameterization functions of the model parameters for a RBM with one visible and one hidden node.", fig.height = 4, fig.align='center', message=FALSE, cache=TRUE}
possibles <- stats(1, 1, type = "binary")

>>>>

We explore the relationship between magnitudes of $\boldsymbol \theta$ and the three measures of model impropriety, near-degeneracy, instability, and uninterpretability, for RBM models of varying size through simulated data.

## Tiny example
To illustrate the ideas of graph near-degeneracy, instability, and uninterpretability in a RBM, we consider first the smallest possible (toy) example that consists of one visible node $v_1$ and one hidden node $h_1$, both binary. A schematic of this model can be found in figure \ref{fig:toymodel}.

\begin{figure}[ht]
  \centering
  \resizebox{1cm}{!}{\input{tikz/toymodel.tikz}}
  \caption{A small example restricted Boltzmann machine (RBM), with two nodes, one hidden and one visible.}
  \label{fig:toymodel}
\end{figure}

@@

Additionally, we can compute the mean parametrization of the model parameters as 
\begin{align*}
\mu(\boldsymbol \theta) &= \text{E}_{\boldsymbol \theta}\left[ t(X) \right] \\
&= \sum\limits_{x \in \mathcal{X}} \left\{ t(x) \frac{\exp\left( \theta_{11} h_1 v_1 + \theta_{h1} h_1 + \theta_{v1} v_1\right)}{\sum\limits_{\boldsymbol x \in \mathcal{X}}\exp\left( \theta_{11} h_1 v_1 + \theta_{h1} h_1 + \theta_{v1} v_1\right)} \right\} \\
&= \sum\limits_{x \in \mathcal{X}} \left\{ t(x) \frac{\exp\left( \theta_{11} h_1 v_1 + \theta_{h1} h_1 + \theta_{v1} v_1\right)}{\gamma(\boldsymbol \theta)} \right\} \\
&= \left[
\begin{matrix}
\frac{\exp\left(\theta_{v_1}\right) + \exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\frac{\exp\left(\theta_{h_1}\right) + \exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\frac{\exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\end{matrix} \right]
\end{align*}
These three functions can be visualized over multiple values of $\boldsymbol \theta$, as seen in figure \ref{fig:degen_toy}. By examining these functions, we can see that as values of $\boldsymbol \theta$ grow larger in magnitude, at least one mean-parameterization functions is approaching the values 0 or 1, indicating a potential to be near to the boundary of the convex hull of $\{t_k(\boldsymbol x): \boldsymbol x \in \mathcal{X}, k = 1, \dots, K\}$. Thus, for a very small example we can see the relationship between values of $\boldsymbol \theta$ and model degeneracy.

```{r near_dependency, fig.cap="\\label{fig:degen_toy}The three mean-parameterization functions of the model parameters for a RBM with one visible and one hidden node.", fig.height = 4, fig.align='center', message=FALSE, cache=TRUE}
possibles <- stats(1, 1, type = "binary")
